Unbiased pangenome graphs

### 2022-10-08 manuscript/main.tex
fix the references

\section{Introduction}
\label{sec:introduction}
A pangenome models the full genomic information of a species or clade \citep{Medini_2005,Sherman_2020}.
In contrast to reference-based approaches that relate sequences to a particular reference genome, methods that use pangenome reference systems attempt to model the mutual relationship between all represented genomes \citep{cpang2018}.
%These models encode the mutual relationships between all the genomes represented, in contrast to reference-based approaches which relate sequences only to a particular ``reference'' genome.
%In principle, pangenome graphs can allow for a great reduction in reference bias, allowing any part of a pangenome to be considered in biological studies. %, and may provide a simpler representation of structurally-variable regions.
Many approaches model the pangenome alignment as a \textit{pangenome graph} \citep{Garrison_2018,Yokoyama2019,Hickey:2020}.
A pangenome graph encodes DNA sequences as walks through an underlying language encoded in a sequence graph \citep{Hein_1989}.
In a pangenome graph, variation can be understood in the context of any part of any included genome  \citep{Eizenga_2020}.
This lets us avoid the problem of reference bias, which can be understood as the limitation of analyses to genome sequences that are similar to a chosen reference genome.
%Reference bias distorts an inferred genome to appear more similar to the reference genome than it actually is.

@@

\begin{definition}
\label{def:vg}
Variation graphs are a common formalism to encode pangenome graphs \citep{Garrison_2019_thesis}.
In the variation graph $\mathcal{G} = (\mathcal{V}, \mathcal{E}, \mathcal{P})$, nodes (or vertices) $\mathcal{V} = v_1\ldots v_{|\mathcal{V}|}$ contain sequences of DNA.
Each node $v_i$ has a unique identifier $i$ and an implicit reverse complement $\bar{v_i}$.
A node strand $s$ corresponds to one node orientation.
Edges $\mathcal{E} = e_1\ldots e_{|\mathcal{E}|}$ connect ordered pairs of node strands ($e_i = ( s_a, s_b )$), encoding the base topology of the graph.
Paths $\mathcal{P} = p_1\ldots p_{|\mathcal{P}|}$ describe walks over node strands ($p_i = s_1 \ldots s_{|p_i|}$), representing the collection of genomes embedded in the graph.
\end{definition}

>>>>

\section{Introduction}
\label{sec:introduction}
A pangenome models the full genomic information of a species or clade \citep{Medini_2005,Sherman_2020}.Garrison_2018_thesis
In contrast to reference-based approaches that relate sequences to a particular reference genome, methods that use pangenome reference systems attempt to model the mutual relationship between all represented genomes \citep{cpang2018}.
%These models encode the mutual relationships between all the genomes represented, in contrast to reference-based approaches which relate sequences only to a particular ``reference'' genome.
%In principle, pangenome graphs can allow for a great reduction in reference bias, allowing any part of a pangenome to be considered in biological studies. %, and may provide a simpler representation of structurally-variable regions.
Many approaches model the pangenome alignment as a \textit{pangenome graph} \citep{Garrison_2018,Yokoyama2019,Hickey:2020}.
A pangenome graph encodes DNA sequences as walks through an underlying language encoded in a sequence graph \citep{Hein_1989}.
In a pangenome graph, variation can be understood in the context of any part of any included genome  \citep{Eizenga_2020}.
This lets us avoid the problem of reference bias, which can be understood as the limitation of analyses to genome sequences that are similar to a chosen reference genome.
%Reference bias distorts an inferred genome to appear more similar to the reference genome than it actually is.

@@

\begin{definition}
\label{def:vg}
Variation graphs are a common formalism to encode pangenome graphs \citep{Garrison_2018_thesis}.
In the variation graph $\mathcal{G} = (\mathcal{V}, \mathcal{E}, \mathcal{P})$, nodes (or vertices) $\mathcal{V} = v_1\ldots v_{|\mathcal{V}|}$ contain sequences of DNA.
Each node $v_i$ has a unique identifier $i$ and an implicit reverse complement $\bar{v_i}$.
A node strand $s$ corresponds to one node orientation.
Edges $\mathcal{E} = e_1\ldots e_{|\mathcal{E}|}$ connect ordered pairs of node strands ($e_i = ( s_a, s_b )$), encoding the base topology of the graph.
Paths $\mathcal{P} = p_1\ldots p_{|\mathcal{P}|}$ describe walks over node strands ($p_i = s_1 \ldots s_{|p_i|}$), representing the collection of genomes embedded in the graph.
\end{definition}
